---
title: "explanation"
date: "September 21, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("bayesplot","knitr","arm","ggplot2","rstanarm","loo","tidyverse",
               "tidyr","dplyr")
```

### MA615 FALL 2021

### HW#2

  - Wrangling
  - Visualization
  
#### Wrangling

This is the first part of this homework, called wrangling.
The data sets are from the Gapminder website (https://www.gapminder.org/data/).
Pick two individual indicators and download each of their data sets into .cvs 
data files. Then, wrangle these two data files into one tibble that is "tidy".

I choose the adult literacy rate and adult employment rate to find the 
relationship between this two indicators. 

Literacy: Adult literacy rate is the percentage of people ages 15 and above who can, 
with understanding, read and write a short, simple statement on their 
everyday life.

Employment: Percentage of total population, age group 15+, that has been employed 
during the given year.

 
##### Import datasets

```{r}
## upload two individual indicators

Literacy_org <- read.csv("/Users/odd/Desktop/FALL2021/MA615/MA615-HW-2/data/literacy_rate_adult_total_percent_of_people_ages_15_and_above.csv", header = T)

employment_org <- read.csv("/Users/odd/Desktop/FALL2021/MA615/MA615-HW-2/data/aged_15plus_employment_rate_percent.csv", header = T)
```

```{r}
## select the adult literacy rate and adult employment rate
Literacy1 <- Literacy_org[, c(1,28)]
sum(is.na(Literacy1$X2000))       # see how many NAs are on the dataset

Literacy_final <- na.omit(Literacy1)
Literacy_final$literacy <- Literacy_final$X2000
# list the literacy rate in 2000 for each country

literacy <- Literacy_final[, c(1,3)]    # divide 
head(literacy)      # look the first 6 rows of the data frame


employment_final <- employment_org[, c(1,12)]
sum(is.na(employment_final$X2000))       # see how many NAs are on the dataset

employment_final$employment <- employment_final$X2000
# list the employment rate in 2000 for each country

Employment <- employment_final[, c(1,3)]
head(Employment)       # look the first 6 rows of the data frame

```


##### Create tibbles
```{r}
## create tibbles
as_tibble(literacy)
as_tibble(Employment)
```


##### Combine two tibbles
```{r}
## combine two tibbles
trend_org <- left_join(literacy, Employment)
trend <- na.omit(trend_org)
trend       # look at the tibble "trend"
```




#### Creating plot

Before we do the visualization, we can guess whether the higher the adult 
literacy rate is, the higher the adult employment rate will be.
Since the majority of jobs require people understanding, read and write a 
short, simple statement, which is the basic requirement, this guess is reasonable.

Now, Create a ggplot visualization of this data.

```{r}
# library packages
library(ggplot2)
library(tidyverse)
library(gapminder)

# library(printr) 
library(RColorBrewer) ## to chose different colors for the graph
```

##### ggplot

Let’s turn this code into a reusable template for making graphs with ggplot2. 
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

```{r}
## To plot trend, run this code to put literacy on the x-axis and 
## Employment on the y-axis
ggplot(data = trend) +
  geom_point(mapping = aes(x = literacy, y = employment))
## The plot shows a positive relationship between literacy rate (literacy) and 
## employment rate (employment) with each country. In other words, high literacy 
## rate leads high employment rate.
```
```{r}
ggplot(data = trend) +
  geom_point(mapping = aes(x = literacy, y = employment, color = country))
## map the colors of points to the class variable to reveal the relationship 
## between literacy rate (literacy) and employment rate (employment) 
## for each country.
```
```{r}
## exploring continuous variales  -- distributions
ggplot(data=trend, aes(x=literacy)) + 
  geom_density()

ggplot(data=trend, aes(x=literacy)) + 
  geom_density(size=1.5, fill="pink", alpha=0.3)

ggplot(data=trend, aes(x=literacy)) + 
  geom_density(size=1.5, fill="pink", alpha=0.5) +
  geom_histogram(aes(y=..density..), binwidth=4, color="black", fill="lightblue", alpha=0.5)

geom_histogram(aes(y=..density..), binwidth=4, color="black", fill="lightblue", alpha=0.5)
## In the majority of those countries, the employment rate has a positive 
## relationship with the literacy rate. We should consider that the first 
## requirment to get a job is recognize the words.
```
```{r}
##  using layers

plt <- ggplot(data=trend,
              aes(x=literacy, y=employment))
plt

plt + geom_point()

plt + geom_point(aes(color=country))

plt + geom_point(mapping = aes(color=country)) +
  geom_smooth(method="loess")

plt + geom_point(aes(color=country)) +
  geom_smooth(mapping = aes(x=literacy, y=employment), method="loess") +
  scale_x_log10()
## use these mappings to extend or overwrite the global mappings
```

##### World Map

```{r, error=FALSE}
## Loading required package: sp
library(sp)
library(lattice)
library(survival)
library(Formula)

library(dplyr)
library(rworldmap) ## plotting the data on World Map
library(countrycode) ## Converting the country name to Country code
library(Hmisc)
```

```{r}
## view trend
dim(trend)

colnames(trend)

sum(complete.cases(trend)) ## No missing values found

describe(trend)  ## see Hmisc
```

```{r}
trend$countrycode <- countrycode(trend$country, 'country.name', 'iso3c')

sPDF <- joinCountryData2Map(trend 
                            ,joinCode = "ISO3"
                            ,nameJoinColumn = "countrycode"
                            ,suggestForFailedCodes = FALSE
                            , verbose = T)

```

```{r}
## Finally, we build the "literacy rate" map over the 29 countries.
colourPalette <- brewer.pal(7,'GnBu')

mapParams <- mapCountryData(sPDF,
                            nameColumnToPlot="literacy",
                            addLegend=FALSE,
                            colourPalette=colourPalette )

## draw a color standard line to displays the literacy rate values 
## corresponding to different colors 

do.call(addMapLegend
        ,c(mapParams
           ,legendLabels="all"
           ,legendWidth=0.5
           ,legendIntervals="data"
           ,legendMar = 2))
```

```{r}
## Finally, we build the "employment rate" map over the 29 countries.
colourPalette <- brewer.pal(7,'GnBu')

mapParams1 <- mapCountryData(sPDF,
                            nameColumnToPlot="employment",
                            addLegend=FALSE,
                            colourPalette=colourPalette )

## draw a color standard line to displays the employment rate values 
## corresponding to different colors 
do.call(addMapLegend
        ,c(mapParams1
           ,legendLabels="all"
           ,legendWidth=0.5
           ,legendIntervals="data"
           ,legendMar = 2))

```

